Comparing tool variability and numerical variability in fMRI analyses

### 2021-10-17 paper/paper.tex
Update result and discussion sections

To measure between-tool (BT) variability, we executed the pipelines described in~\cite{bowring2019exploring}
with FSL version 5.0.10, AFNI version 18.1.09, and SPM12 version r7771
executed with GNU/Octave version 5.2.
All these software versions were identical to the original study except for SPM
that we used Octave instead of MATLAB to enable mathematical function instrumentation using FL,
while MATLAB uses its built-in functions and does not allow to apply FL.
We also used a fixed number of 7 threads in AFNI executions,
to reproduce similar results in~\cite{bowring2019exploring} in a feasible execution time.
All the analyses were conducted on the same operating system, CentOS 7.3.
%We ensured that the software versions \tristan{you mentioned versions before, the writing is a bit circular here} used in all experiments were similar to the original study,
We encapsulated the above-mentioned software packages in Docker container images with available Dockerfiles at \url{https://github.com/ali4006/fuzzy-neurotools/tree/main/dockerfile}.

@@

% \tristan{did you check standard error vs standard deviation?}\ali{STD shows variability of data in relation to the mean (what we are looking for here),
% while SE is a type of STD (computed as $\frac{\sigma}{\sqrt{n}}$) for distribution of means (we use it when we are interested in the precision of means in different samples).}

@@

We ensured a successful replication of the analyses initially presented in~\cite{bowring2019exploring}
by comparing t-statistic group maps with results in the original study.
We found identical checksums in the results of FSL.
Also, we could replicate results with the same SPM version and MATLAB as in the original study with visually imperceptible differences.
Changes in the SPM version and replacing Octave with MATLAB resulted in visually slight differences which were still hard to distinguish between them.
For AFNI, the activation maps were similar overall; however, visual differences were noticeable.
This might be caused by the hardware optimizations such as dynamic scheduling~\cite{demmel2013numerical}.
We could confirm that AFNI results were passed QC by visually assessing the registration results of each individual subject.

Summary statistics for the group-level unthresholded and thresholded t-statistics are reported in Table~\ref{table:pipeline-stats}.
% The variability between results of each pair of tools in BT and individual tools in WT was measured using the mean and standard deviation of absolute differences.
Overall, WT variability was an order of magnitude lower than BT variability for both means and standard deviations.
While the highest variability was observed in \fslafni, \fslspm was made the least variations in BT.
Results also show that AFNI was generated the most variability in WT among the three tools \tristan{this needs statistical testing}.

@@

% Fig2
Figure~\ref{fig:dice-thresh} compares regional normalized Dice coefficients of
activated voxels, showing a linear correlation between BT and WT with p-value $< 0.05$ in all three pairs.
% The vertical line where the Dice score is zero in BT demonstrates regions in the brain with no overlaps between activated voxels in BT but WT,
% including left/right `V1\_ROI'\footnote{The primary visual cortex} in \fslafni and \fslspm,
% and left/right `TGd\_ROI'\footnote{The area TG dorsal (part of lateral temporal cortex)} in \afnispm results.

@@

  %%%%%%%%%% Corr. plot of tstats%%%%%%%%
  \begin{figure*}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std-corr-unthresh-plot.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \hfill
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std/correlated-unthresh.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \caption{Scatter plots of unthresholded group t-statistics standard deviations computed between tools and within tools.
        Identity lines distinguish voxels where the standard deviation ratio of BT/WT is in the interval [0.5, 2]
        and are represented on brain maps in the second row.}
    \label{fig:unthresh-correlation}
    \end{minipage}}
  \end{figure*}

@@

We determined the virtual precision of t=17 bits as the precision that minimized the RMSE between BT and WT on average.
This is the precision that numerical perturbations in FSL more closely simulated BT variability.
The brain maps of standard deviations between BT and WT obtained from the Fuzzy FSL at t=17 bits (Figure~\ref{fig:gnp-mni})
show significant WT variability, particularly for some regions in the frontal, parietal, and limbic lobes, close to the BT variability.

@@

  \item[$\bullet$ ] In this study, we represented the magnitude of differences in BT and WT results.
  We obtained more instability in BT compared to WT. Also we showed how between tool variations
  are correlated with numerical variability.
  While we only perturbed basic mathematical functions, perturbing linear algebra, in particular, could increase WT variability.

  \item[$\bullet$ ] Generally, we obtained more uncertainty on thresholded results, probably due to different thresholding
  methods used in different tools. This can raise further investigations on the thresholding techniques toward stability.
  Also, we obtained significant instabilities in specific subjects so that healed by averaging in group-level analyses.

  \item[$\bullet$ ] We showed that reducing the number of virtual precision bits leads to more instabnilities in which can exceeds variability compared to BT.
  We also showed the numerical perturbations can perfectly simulate BT variability in some regions.

  \item[$\bullet$] Numerical instability in individual analyses is attenuated in group analyses. This is an issue for the development of fMRI biomarkers.

  \item[$\bullet$] Comparision between FSL, SPM and AFNI. Discussion on the instrumentation of math functions only.
\end{itemize}

>>>>

To measure between-tool (BT) variability, we executed the pipelines described in~\cite{bowring2019exploring}
with FSL version 5.0.10, AFNI version 18.1.09, and SPM12 version r7771
executed with GNU/Octave version 5.2.
All these software versions were identical to the original study except for SPM
that we used Octave instead of MATLAB to enable mathematical function instrumentation using FL,
while MATLAB uses its built-in functions and does not allow to apply FL.
We also used a fixed number of 7 threads in AFNI executions,
to reproduce similar results in~\cite{bowring2019exploring} in a feasible execution time.
All the analyses were conducted on the same operating system, CentOS 7.3.
We encapsulated the above-mentioned software packages in Docker container images with available Dockerfiles at \url{https://github.com/ali4006/fuzzy-neurotools/tree/main/dockerfile}.

@@

@@

We ensured a successful replication of the analyses initially presented in~\cite{bowring2019exploring}
by comparing t-statistic group maps with results in the original study.
We found identical checksums in the results of FSL.
Also, we could replicate results with the same SPM version and MATLAB as in the original study with visually imperceptible differences.
Changes in the SPM version and replacing Octave with MATLAB resulted in visually small differences which were still hard to distinguish between them.
For AFNI, the activation maps were similar overall; however, visual differences were noticeable.
This might be caused by the hardware optimizations such as dynamic scheduling~\cite{demmel2013numerical}.
We could confirm that AFNI results were passed the quality checks by visually assessing the registration results of each individual subject.

Summary statistics for the group-level unthresholded and thresholded t-statistics are reported in Table~\ref{table:pipeline-stats}.
Overall, WT variability was an order of magnitude lower than BT variability for both means and standard deviations.
While the highest variability was observed in \fslafni, \fslspm was made the least variations in BT.
We also computed t-test statistics to examine differences in WT variability between tools.
Results show that AFNI was generated the most variability in WT among the three tools with 
the high t-value 39.6 between \fslafni samples and 65.5 between \afnispm samples, and p-value $< 0.05$.
\ali{I'm checking this with Greg}.

@@

% Fig2
Figure~\ref{fig:dice-thresh} compares regional normalized Dice coefficients of
activated voxels, showing a linear correlation between BT and WT with p-value $< 0.05$ in all three pairs.

@@

  %%%%%%%%%% Corr. plot of tstats%%%%%%%%
  \begin{figure*}[ht]
    \fbox{\begin{minipage}{\dimexpr \textwidth-2\fboxsep-2\fboxrule}
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std-corr-unthresh-plot.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \hfill
      \begin{subfigure}[ht]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/std/correlated-unthresh.png}
        %\caption{Standard deviation of thresholded t-statistics map on template surface}
      \end{subfigure}
      \caption{Scatter plots of unthresholded group t-statistics standard deviations computed between tools and within tools.
        Identity lines distinguish voxels where the standard deviation ratio of BT/WT is in the interval (0.5, 2)
        and are represented on brain maps in the second row.}
    \label{fig:unthresh-correlation}
    \end{minipage}}
  \end{figure*}

@@

We determined the virtual precision of t=17 bits as the precision that minimized the RMSE between BT and WT on average.
This is the precision that numerical perturbations in FSL more closely simulated BT variability.
The brain maps of standard deviations between BT and WT obtained from the Fuzzy FSL at t=17 bits (Figure~\ref{fig:gnp-mni})
show significant WT variability, particularly for some regions in the frontal, parietal, and limbic lobes, close to BT variability.

@@

We represented differences in the results of an fMRI analysis between numerical variability
and the variability arisen from software package variations.
Although BT variability was an order of magnitude higher than WT variability on average,
we demonstrated regions in the brain maps that numerical perturbations could simulate BT variability.
The correlation of these variabilities suggests that numerical perturbations can be used to evaluate the robustness of the analyses
to inter-software disparities.

While switching between \fslspm was generated the least BT and WT variabilities consistently, the most level of these variabilities was identified
between \fslafni and \afnispm. We also found AFNI as the most sensitive tool to numerical perturbations.
These results do not indicate which software package is better or worse since there is no ground-truth model for the analyses.
However, this study revealed the importance of numerical instability in the selected tools for doing fMRI analyses.
It provided a view of related numerical variability of which should be taken into consideration by the software developers.

We found that numerical instability in individual analyses was attenuated in group analyses
since we obtained substantially higher variability in t-statistics at the level of subjects.
This is consistent with the findings in~\cite{bowring2021isolating} in which identified the first-level signal model as the largest
source of variability across the three tools for this study.
This is an issue for the development of fMRI biomarkers.
As a solution, we can aggregate the results obtained from numerical perturbations to improve biomarkers,
as suggested in~\cite{kiar2020numerical} where data augmentation using numerical perturbations resulted in a better classification in connectomics.

It is notable that we obtained more uncertainty on thresholded results than unthresholded maps, %This issue is 
probably due to different thresholding values used in different tools.
This is consistent with the observations in~\cite{bowring2021isolating} where identified a high correlation between unthresholded t-statistic maps,
but substantially low Dice coefficient for activations in the thresholded maps.
This raises further investigations on thresholding techniques to reduce instabilities.

Given that FL only perturbs basic mathematical functions, we expect more numerical variability by perturbing linear algebra,
in particular for non-scalar types like vectors or matrices. For the future study, this can be evaluated using MCA-instrumented versions 
of BLAS and LABACK along with other libraries available in the Fuzzy project in Verificarlo's GitHub repository at github.com/verificarlo/fuzzy.
